---
title: 'Chapter 2 Lab: Introduction to R'
author: Gowreesh Gunupati
date: "03-17-2022"
output:
  html_document: default
  pdf_document: default
---

**Instructions and time-saving hints.**  To begin this lab, download the archive and unpack it.  Inside the folder, you will find a Rmarkdown file and csv files for lab data.  Please keep the Rmarkdown file in the same directory as the data files.

The purpose of the first part of this lab is to quickly review the basics of R.  To review the basics, you will reproduce the commands blocks in JW Section 2.3 in this Rmarkdown file.  **Please do not retype all the R commands as shown in Section 2.3.**  Instead, go to [link to text for Lab 2](http://faculty.marshall.usc.edu/gareth-james/ISL/Chapter%202%20Lab.txt), cut and paste the text into your Rmarkdown file.  Then break up the R commands into the R chunks shown in the text.  (R chunks are braced by triple backticks and a leading {r}, as before.)  Do not put all commands into the same block!  The idea is to imitate the code chunks in Section 2.3 with the additional plots and output omitted from the text.  Below, I have given you the first two blocks for reference.

After breaking the text into the correct chunks, you may knit the document.
You will notice two problems when knitting.  The first is the appearance of a data editor window (3 times), due to the `fix()` commands in the blocks.  You may simply quit these windows each time, or comment out the `fix()` statements.  The second problem is an error message, due to the block at the top of p. 50 in JW.  You may comment out this `plot()` statement to knit again.

Congratulations!  When the document is successfully knitted, you should read Section 2.3 along with the Rmarkdown output.  You should see the commands along with the output plots.  Of course, the blocks with help commands, fix() commands, and intentional errors will not be reproduced.  Together, JW2.3 and your Rmarkdown file will help you to review some basics of R.  Note that the plotting is done with core R commands and not with the `ggplot2` package as in Homework 0.

 Next, you will apply these core R basics to the `College.csv` dataset, as described in problem 8, JW p.54.  Please complete the Rmarkdown document corresponding to problem statement.  Submit your Rmd file along with an unzipped PDF of the result before the deadline.
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic Commands


```{r}
x <- c(1,3,2,5)
x
```

```{r}
x = c(1,6,2)
x
y = c(1,4,3)
# Chapter 2 Lab: Introduction to R
```



# Basic Commands
```{r}
?ls()
```

```{r}
?rm()
```

```{r}
x <- c(1,3,2,5)
x
x = c(1,6,2)
x
y = c(1,4,3)
length(x)
length(y)
x+y
ls()
rm(x,y)
ls()
rm(list=ls())
```


```{r}
?matrix
x=matrix(data=c(1,2,3,4), nrow=2, ncol=2)
x
```


```{r}
x=matrix(c(1,2,3,4),2,2)
```


```{r}
matrix(c(1,2,3,4),2,2,byrow=TRUE)
```


```{r}
sqrt(x)
```


```{r}
x^2
```


```{r}
x=rnorm(50)
y=x+rnorm(50,mean=50,sd=.1)
cor(x,y)
```


```{r}
set.seed(1303)
rnorm(50)
set.seed(3)
y=rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)


```

# Graphics
```{r}
?dev.off()
```

```{r}
x=rnorm(100)
y=rnorm(100)
plot(x,y)
plot(x,y,xlab="this is the x-axis",ylab="this is the y-axis",main="Plot of X vs Y")
pdf("Figure.pdf")
plot(x,y,col="green")
```


```{r}
dev.off()
x=seq(1,10)
x
x=1:10
x
x=seq(-pi,pi,length=50)
y=x
f=outer(x,y,function(x,y)cos(y)/(1+x^2))
contour(x,y,f)
contour(x,y,f,nlevels=45,add=T)
fa=(f-t(f))/2
contour(x,y,fa,nlevels=15)
image(x,y,fa)
persp(x,y,fa)
persp(x,y,fa,theta=30)
persp(x,y,fa,theta=30,phi=20)
persp(x,y,fa,theta=30,phi=70)
persp(x,y,fa,theta=30,phi=40)


```

# Indexing Data
```{r}
A=matrix(1:16,4,4)
A
A[2,3]
```


```{r}
A[c(1,3),c(2,4)]
```


```{r}
A[1:3,2:4]
```


```{r}
A[1:2,]
A[,1:2]
A[1,]
A[-c(1,3),]
A[-c(1,3),-c(1,3,4)]
dim(A)


```

# Loading Data

```{r}
Auto=read.table("Auto.data")
```

```{r}
?fix()
```

```{r}
?fix(Auto)
```


```{r}
Auto=read.table("Auto.data",header=T,na.strings="?")
#fix(Auto)
Auto=read.csv("Auto.csv",header=T,na.strings="?")
#fix(Auto)
dim(Auto)
Auto[1:4,]
Auto=na.omit(Auto)
dim(Auto)
names(Auto)
```


```{r}
?mpg
```


# Additional Graphical and Numerical Summaries
```{r}
#plot(cylinders,mpg)
plot(Auto$cylinders, Auto$mpg)
attach(Auto)
plot(cylinders, mpg)
cylinders=as.factor(cylinders)
plot(cylinders, mpg)
plot(cylinders, mpg, col="red")
plot(cylinders, mpg, col="red", varwidth=T)
plot(cylinders, mpg, col="red", varwidth=T,horizontal=T)
plot(cylinders, mpg, col="red", varwidth=T, xlab="cylinders", ylab="MPG")
hist(mpg)
hist(mpg,col=2)
hist(mpg,col=2,breaks=15)
#pairs(Auto)
pairs(~ mpg + displacement + horsepower + weight + acceleration, Auto)
plot(horsepower,mpg)
identify(horsepower,mpg,name)
summary(Auto)
summary(mpg)
```


#Applied JW p.54 (8.)##

Here, we are using the `College` data set, found in  `College.csv`. 
Before reading the data into R, it can be viewed in Excel or a text editor.

1. Use the `read.csv()` function to read the data into R. Call the loaded data `college`. Make sure that you have the directory set to the correct location for the data.

```{r}
college<-read.csv("College.csv",header=TRUE)
head(college)
```

2. Look at the data using the `fix()` function. You should notice that the first column is just the name of each university. We don’t really want R to treat this as data. However, it may be handy to have these names for later. Try the following commands:

```{r}
rownames(college)=college[,1]
```


```{r}
rownames
```


```{r}
head(college)
```


```{r}
#fix(college)
```

You should see that there is now a `row.names` column with the name of each university recorded. This means that R has given each row a name corresponding to the appropriate university. R will not try to perform calculations on the row names. However, we still need to eliminate the first column in the data where the names are stored. Try

```{r}
college=college[,-1] 
#fix(college)
```

Now you should see that the first data column is `Private`. Note that another column labeled `row.names` now appears before the `Private` column. However, this is not a data column but rather the name that R is giving to each row.


3.   Please complete these parts.

    Use the `summary()` function to produce a numerical summary of the variables in the data set.  
    
```{r}

?college
summary(college)
```

Use the `pairs()` function to produce a scatterplot matrix of the first ten columns or variables of the data. Recall that you can reference the first ten columns of a matrix `A` using `A[,1:10]`.  
    
```{r}
pairs(college[,sapply(college[,1:10],is.numeric)])
```

Use the `plot()` function to produce side-by-side boxplots of `Outstate` versus `Private`.
    
```{r}
?plot()

boxplot(college$Outstate~college$Private,xlab="Status",ylab="Outstate",col=c("red","blue"),names=c("Public","Private"))

```

Create a new qualitative variable, called `Elite`, by *binning* the `Top10perc` variable. We are going to divide universities into two groups based on whether or not the proportion of students coming from the top 10% of their high school classes exceeds 50 %.
    
```{r}
?as.factor()
Elite=rep("No",nrow(college))
Elite[college$Top10perc >50]="Yes"
Elite=as.factor(Elite)
college=data.frame(college ,Elite)

```

Use the `summary()` function to see how many elite universities there are. Now use the `plot()` function to produce side-by-side boxplots of `Outstate` versus `Elite`.
    
```{r}
summary(Elite)
boxplot(Outstate ~ Elite, data=college,xlab="Elite Status",ylab="Outstate",main="Outstate By Elite Status",col=c("red","yellow"),names=c("Elite","Other Colleges"))
```
Use the `hist()` function to produce some histograms with 5 and 50 bins for `Outstate` and `Apps`. Use the command `par(mfrow=c(2,2))`: it will divide the print window into four regions so that four plots can be made simultaneously. 
    
```{r}
par(mfrow=c(2,2))
hist(college$Outstate,breaks=5)
hist(college$Outstate,breaks=50)
hist(college$Apps,breaks=5)
hist(college$Apps,breaks=50)
```

*Continue exploring the data, and provide a brief summary of what you discover.* 

This statement is intentionally vague: as a data scientist, it will be your job to propose hypotheses about the data, and then to use the data to address your hypotheses.  This is a creative, iterative, (and fun) process. 

In this first lab I will propose four hypotheses for you.  You must address these hypotheses with the suggested approaches. The approach will be to produce one plot for each hypothesis in a 2 by 2 set of plots.  Discuss your findings for each case, including: whether the data supports or rejects the hypothesis, and to what degree the hypothesis is rejected or supported.  After addressing the four hypotheses, I would like for you to propose at least one more hypothesis, and  develop a methodology.  Brevity matters, and less is best.  Produce a single plot upon which to base your answer.  Long answers will be given little credit.  For this last part, you will be graded on the "interest" of the hypothesis, your approach to address it, and your discussion.  

    
*Hypothesis 1. The tuition at the best colleges, as indicated by `Elite`, is higher than that at other colleges.*  **Methodology:** Use one-over-other horizontal red boxplots of `Outstate` for elite and non-elite colleges.  Title and label the axes of this plot thoughtfully. Carefully use this single plot to address this hypothesis.  Can you be precise about the word *higher* ?

```{r}
attach(college)
#Hypothesis 1 <- 
boxplot(Outstate ~ Elite, horizontal = TRUE, col = "red",
        xlab = "Outstate Tution fee", ylab = "Elite Status", main = "Tuition fees of Outstate")

#Observation: We can say that the Hypothesis is supported by the by the boxplot below we can say that The tuition at the best colleges, as indicated by `Elite`, is higher than that at other colleges
```

The enrollment rate is the fraction of accepted students who enrolled.
*Hypothesis 2. The enrollment rate at the best colleges, as indicated by `Elite`, is higher than that at other colleges.*  **Methodology:** Create a new variable called `EnrollRate`, using `Enroll` and `Accept`.  Use the `attach()` function to make your code cleaner. Use one-over-other horizontal green boxplots of `EnrollRate` for elite and non-elite colleges.  Title and label the axes of this plot thoughtfully. Carefully use this single plot to address this hypothesis.  Can you be precise about the word *higher* ? Can you comment on the presence of outliers in the boxplots?
```{r}
#Hypothesis 2 

#Created a new variable called `EnrollRate`, using `Enroll` and `Accept

EnrollRate <- Enroll /Accept

# Create a one-over-other horizontal green boxplot of EnrollRate for elite and non-elite colleges

boxplot(EnrollRate ~ Elite,horizontal=TRUE, col="green",ylab="Elite Status", xlab="Enrollment Rate",main="Enrollment Rate by Elite Status")

#Observation: By the boxplot we can say that the results doesn't support the Hypothesis a little as we can see a slightly lower median of enrollment rate for the ELite colleges than that of other colleges

#Can you comment on the presence of outliers in the boxplots?
# The presence of outliers in the hyposthesis 2 plot states that few colleges which are not elite the Enrollment rate is greater than 0.8 and also some colleges have the extreme outlier of having the enrollment rate as 1. This can be seen by the points or bubbles after the third quartilein the boxplot.
```


*Hypothesis 3. The number of applications per enrolled student is higher at elite colleges than other colleges.* **Methodology:** Use one-over-other horizontal blue boxplots of applications per enrolled student for elite and non-elite colleges. Title and label the axes of this plot thoughtfully. Carefully use this single plot to address this hypothesis.  Can you be precise about the word *higher* ?  Can you comment on the presence of outliers in the boxplots?

```{r}
apps_by_enrollstudent<-Apps/Enroll

#Hypothesis 3:
boxplot(apps_by_enrollstudent ~ Elite, horizontal = T, col = "blue",
        main = "Applications per Enrolled Student by Elite Status", xlab = "applications for Elite colleges per enrolled student", ylab = "Elite")

#Obseration- By the Boxplot we can say that the result supports the hypothesis the number of applications per enrolled student is higher at elite colleges than other colleges.

#Can you comment on the presence of outliers in the boxplots?
# The presence of outliers in the hyposthesis 3 plot states that there are few colleges which are not elite got more applications per enrolled rate and even it went till 13-14 applications per enrolled student.As mentioned these colleges are few but the applications recieved per enrolled student is more and this can be proved using the box plot down
```
*Hypothesis 4.  The fraction of alumni who donate is higher at elite colleges than other colleges.* **Methodology:** Use one-over-other horizontal cyan boxplots of `perc.alumni` for elite and non-elite colleges. Title and label the axes of this plot thoughtfully. Carefully use this single plot to address this hypothesis.  Can you be precise about the word *higher* ?

```{r}
#Hypothesis 4:
  boxplot(perc.alumni ~ Elite, horizontal = TRUE, col = "cyan",
        main = "Alumni Donation by Elite Status", xlab = "Alumni Donation Rate", ylab = "Elite Status")

#Observation: By seeing the Boxplot we can say that result is supporting the Hypothesis The fraction of alumni who donate is higher at elite colleges than other colleges
```
*Hypothesis 5.  The graduation rate is high in elite colleges than other colleges.* **Methodology:** Use one-over-other horizontal orange boxplots of `Grad.rate` for elite and non-elite colleges. Title and label the axes of this plot thoughtfully. Carefully use this single plot to address this hypothesis.  Can you be precise about the word *higher* ?
 

```{r}
#Hypothesis 5:
 boxplot(Grad.Rate ~ Elite, horizontal = TRUE, col = "orange",
        main = "Graduation rate by Elite Status", xlab = "Graduation Rate", ylab = "Elite Status")

#Observation: By seeing the Boxplot we can say that result is supporting the Hypothesis the graduation rate is high in elite colleges than other colleges
```

*Hypothesis 6.  The Student faculty Ratio is low in elite colleges than other colleges.* **Methodology:** Use one-over-other horizontal pink boxplots of `S.F ratio` for elite and non-elite colleges. Title and label the axes of this plot thoughtfully. Carefully use this single plot to address this hypothesis.  Can you be precise about the word *lower* ?

```{r}
#Hypothesis 6:
 boxplot(S.F.Ratio ~ Elite, horizontal = TRUE, col = "pink",
        main = "Student Faculty Ratio by Elite Status", xlab = "Student faculty ratio", ylab = "Elite Status")


#Observation: By seeing the Boxplot we can say that result is supporting the Hypothesis the Student faculty Ratio is low in elite colleges than other colleges 
```

*Hypothesis 7.  The number of PhD's is high in elite colleges than other colleges.* **Methodology:** Use one-over-other horizontal yellow boxplots of `Phd` for elite and non-elite colleges. Title and label the axes of this plot thoughtfully. Carefully use this single plot to address this hypothesis.  Can you be precise about the word *higher* ?

```{r}
#Hypothesis 7:
 boxplot(PhD ~ Elite, horizontal = TRUE, col = "yellow",
        main = "Number of Phds by Elite Status", xlab = "No of PhDs ", ylab = "Elite Status") 


#Observation: By seeing the Boxplot we can say that result is supporting the Hypothesis the Student faculty Ratio is low in elite colleges than other colleges  
detach(college)
  
```

